Estimating initial state and quark-gluon plasma medium properties \\ using a hybrid model with nucleon substructure \\ calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV

### 2018-07-31 paper/substructure.tex
Partway through summary and conclusion.

\section{summary}
\label{sec:summary}

In this work, we constructed a multi-stage hybrid transport model for relativistic nuclear collisions and calibrated free parameters of the model using Bayesian parameter estimation.
Specifically, we focused on improving the model's description of small collision systems by investigating the effect of nucleon substructure on bulk observables in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.

The analysis was performed using a parametric data-driven approach to account for theoretical uncertainty (variability) in the QGP initial conditions and hydrodynamic transport coefficients using a handful of equally uncertain model parametersi.
We marginalized over the theoretical uncertainty in the parameter using a handful of equally uncertain model parameters to vary the size and shape of the nucleon, the scaling behavior of initial QGP entropy deposition, the duration of pre-equilibrium dynamics, and the temperature dependence of the hydrodynamic shear and bulk viscosities.

We placed a conservative prior on the ``true'' value of each parameter, and calculated a global likelihood

varied each parameter within its design range to explored the full 15-dimensional parameter space of our model.

constructed an emulator to interpolate the predictions of our model at arbitrary regions of parameter space.

Bayesian inference was then used to compare the predictions of the model to experimental data from CMS and ALICE, in order to quantify the posterior probability that a given set of model parameters describes the experimental data

\medskip

>>>>

\section{summary and conclusions}
\label{sec:summary}

Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \note.
This narrative is evidenced by the global, self-consistent and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \note.
Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the so-called dilute limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
Tell-tale signatures of hydrodynamic collectivity were thus always expected to vanish in smaller nuclear collision systems, e.g.\ $p$-$p$ and $p$-Pb collisions, where the number of produced particles is orders of magnitude smaller than a typical Pb-Pb collision.

These expectations were upended, however, when long-range multiparticle correlations were detected in high-multiplicity $p$-Pb collisions and found to be similar in magnitude to those observed in Pb-Pb collisions \note.
Nuclear collision systems which were previously thought to be too small for hydrodynamic flow, were subsequently found to generate the same collectivity used to justify hydrodynamic flow in heavy-ion collisions.
It is thus natural to wonder if a single unified hydrodynamic framework might be able to describe $p$-Pb and Pb-Pb bulk observables simultaneously.

In this work, we performed a semi-exhaustive search for a unified description of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using Bayesian methods to rigorously calibrate and constrain free parameters of a flexible nuclear collision model based on viscous hydrodynamics.
The goal of our study was two fold.
First, we aimed to establish whether or not our hydrodynamic framework was able to describe $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV simultaneously, i.e.\ we attempted to falsify the framework.
And second, given that the former is true, we wished to obtain estimates for the \emph{true} model parameters given the assumptions of our model and the evidence provided by the experimental data.

Perhaps the most important difference between the present study and previous Bayesian analyses of heavy-ion collision models \note\ is the inclusion of variable nucleon substructure.
We described the QGP initial state using a modified version of the parametric \trento\ model, with new parameters to vary the fluctuating size and shape of each nucleon.
The nucleon was modeled as a cluster of $n_c$ Gaussian constituents (hot spots), and each constituent was described by a Gaussian density profile of width $v$.
We also varied a parameter $w$ which controlled the sampling radius of the constituent positions inside each nucleon and hence the resulting RMS width of the ensemble averaged nucleon density.

The transport dynamics were simulated using an event-by-event multi-stage nuclear collision model described by a pre-equilibrium free-streaming stage coupled to boost-invariant viscous hydrodynamics for the early hot and dense regions of the collision, and a microscopic hadronic afterburner to simulate hadronic interactions .
We parametrized various sources of uncertainty including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the switching or particlization temperature where the medium description switches from hydrodynamics to microscopic Boltzmann transport.

%Unfortunately, validating (or invalidating) a hydrodynamic picture of $p$-Pb and Pb-Pb collisions is not as simple as comparing a single model calculation to experimental data.
%Hydrodynamic simulations typically include a number of free parameters which quantify uncertainties in the initial conditions and QGP medium, and these parameters can dramatically affect model predictions.
%It's thus imperative to test all possible combinations of parameters in order to compare the full landscape of physically reasonable model calculations to experimental data.
%If all possible combinations of model parameters fail to describe both collision systems simultaneously, then it's probably safe to say that the underlying assumptions of the chosen framework are either incomplete or invalid.
%Similarly, if a single set of model parameters describes $p$-Pb and Pb-Pb bulk observables simultaneously, then it's compelling evidence that the hydrodynamic framework generalizes to small collision systems.
%
%There is also much more that can be learned about a model by testing all possible parameter combinations as opposed to a single set of values.
%If the model works exceptionally well, but only for a certain range of parameter values, it's evidence that those parameter values are somehow constrained by the data.
%Moreover, rigorous statistical methods may then be used to report parameter estimates with meaningful uncertainties which account for the accuracy of the chosen framework and the resolving power of the experimental data.
